Toronto's Covid-19 Impact: Effect of Age and Gender on Outcomes

### 2024-09-24 paper/paper.qmd
edits to cleaned data, cleaning script, testing script, and paper

---
title: "Toronto's Covid-19 Impact: How did Age and Gender affect Outcomes?"
author: 
  - Dhruv Gupta
thanks: "Code and data are available at: https://github.com/dhruv5423/Covid19-R-Project"
date: "today"
date-format: long
abstract: "The Covid-19 Pandemic caused irreparable damage to the infrastructure of our global systems, highlighting the neccessity to better understand factors that influence health outcomes. This paper uses data from OpenDataToronto to investigate how age and gender affects outcomes of those infected with the virus. xyz about results"
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

#Loading Packages
library(tidyverse)
library(here)
library(palmerpenguins)
library(opendatatoronto)
library(tinytex)

@@

The remainder of this paper is structured as follows. @sec-data xyz @sec-results xyz @sec-discussion xyz.

@@

## Data Selection and Measurement

@@

@@

Talk way more about it.

@@

>>>>

---
title: "Toronto's Covid-19 Impact: Effect of Age and Gender on Outcomes"
author: 
  - Dhruv Gupta
thanks: "Code and data are available at: https://github.com/dhruv5423/Covid19-R-Project"
date: "today"
date-format: long
abstract: "The Covid-19 Pandemic caused irreparable damage to the infrastructure of our global systems, highlighting the neccessity to better understand factors that influence health outcomes. This paper uses data from OpenDataToronto to investigate how age and gender affects outcomes of those infected with the virus. xyz about results"
format: pdf
toc: true
number-sections: true
bibliography: references.bib
header-includes:
  - \usepackage{float}
---

@@

#Loading Packages
library(pander)
library(tidyverse)
library(here)
library(palmerpenguins)
library(opendatatoronto)
library(tinytex)
library(knitr)
library(kableExtra)

@@

The remainder of this paper is structured as follows. @sec-data discusses @sec-results xyz @sec-discussion xyz.

@@

## Data Selection

Data used in this report was sourced from OpenDataToronto's portal. @opendatatoronto. More specifically, the dataset "Covid-19 Cases in Toronto" was used and cleaned for the purposes of this report. Toronto Public Health (TPH) released anonymized, person-level data from the start of the pandemic in January 2020. The data spans from the first reported case on in January of 2020 to February 14th, 2024. In a statement on the website for this dataset, OpenDataToronto states that "As case and outbreak management guidelines changed and COVID-19 specific resources were no longer funded, the level of detail available for cases decreased, and more recent data are less complete and not comparable to previous years. TPH discontinued the production of this report with the final refresh as of February 14, 2024" @opendatatoronto_set

@@

In it's original form, the dataset contains more than 414,000 entries regarding information on cases of Covid-19 in Toronto. Below are two tables with the first three rwos of the raw data, separated into two tables for readability.

### Table 1: Sample of Raw Covid-19 Data

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(kableExtra)

# First part of Table 1 (first few columns)
raw_covid_data %>%
  select(X_id, Outbreak.Associated, Assigned_ID, Age.Group, Neighbourhood.Name, FSA, Source.of.Infection, Classification) %>%
  head(3) %>%
  knitr::kable(booktabs = TRUE) %>%
  kable_styling(font_size = 8)

# Second part of Table 1 (next set of columns)
raw_covid_data %>%
  select(Episode.Date, Reported.Date, Client.Gender, Outcome, Ever.Hospitalized, Ever.in.ICU, Ever.Intubated) %>%
  head(3) %>%
  knitr::kable(booktabs = TRUE) %>%
  kable_styling(font_size = 8)

```

(\scriptsize Table Separated using Tips from [@stack_tip] and LLMs)

@@

### Table 2: Sample of Cleaned Covid-19 Data
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Cleaned Data Table (Table 2)
cleaned_covid_data %>%
  head(5) %>%
  kable(col.names = c("x_id", "age_group", "client_gender", "reported_date", "ever_in_icu", "ever_hospitalized", "ever_intubated", "outcome"),
        booktabs = TRUE)
```

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE}
cleaned_covid_data %>%
  summarise(
    total_cases = n(),
    age_group_distribution = table(age_group),
    gender_distribution = table(client_gender),
    outcome_distribution = table(outcome),
    earliest_reported_date = min(reported_date, na.rm = TRUE),
    latest_reported_date = max(reported_date, na.rm = TRUE)
  )
```
## A Note on Measurement

-   gold data quality
-   updates on funding/measurement changes + discontinuation
-   likely many unreported cases
